### 2019-03-05 src/manuscript/pandoc-frontiers-template/demo-manuscript.Rmd
Set up Frontiers template and created demo manuscript

---
title: 'Demo Manuscript'
author: "An author"
date: "October 30th 2018"
abstract: "Hello, this is the abstract."
keywords: "keyword1, kw2, kw3... have to have at least 5, max 8"
language: en-US
output:
  pdf_document:
    dev: pdf
    keep_tex: true
    template: ./pandoc-frontiers-template.tex
---

Hello, world!

# Introduction
Here are some things to consider:

1. We don't know what's happening
2. Neither does anyone else

Also important:

- this creates an unordered list
- and this is the second item.

This is how links work: [link text](https://example.com/)

We can also include code snippets:
```python
# Examplary python code
print("Hello World!")
print("2 + 5 = " + (2+5))
with open('testfile.fa', 'w') as file:
  file.write("Hello, world")
print("Done")
```

# Rmarkdown
Now for some really kewl stuff: We load a file in R:

```{r loadSomething}
annotation_quantities = 
  read.csv('../../../data/mocks/processed/annotation_quantities.csv', header=T)
rice = subset(annotation_quantities, genome=='rice' & source=='GOMAP')
```

And now we can actually use it:
```{r demo1}
head(annotation_quantities)
summary(annotation_quantities$n_c)
```

The annotation set for rice had a total of `r rice$n_c + rice$n_p + rice$n_f` annotations.

## Plots
We can even plot the data:
```{r plot}
a = annotation_quantities
pie(a$n_c+a$n_f+a$n_p, labels=paste(a$genome, a$source))
```

```{r plot2}
barplot(a$n_c+a$n_f+a$n_p, las=2)
```