### 2019-08-14 text/index.Rmd

---
title: "This is the title"
author: "The Pickles"
date: "Febcember 32nd 3023"
language: en-US
site: "bookdown::bookdown_site"
output:
  bookdown::pdf_book:
    keep_tex: true
    keep_md: true
    includes:
      in_header:
        preamble.tex
documentclass: book
---

```{r load_libs, include=F}
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo=FALSE)
```

# Introduction
Hello, how are we doing?