### 2019-08-30 text/index.Rmd
List of Authors (preliminary)

---
title: "Article Title"
runtitle: "Short Title"
documentclass: frontiersSCNS  # or frontiersHLTH, or frontiersFPHY
author:
  - name: Dennis Psaroudakis
    affiliation: '1'
    etal: Psaroudakis # First author's last name. 
  - name: Ha Vu
    affiliation: '1'
  - name: Colleen Yanarella
    affiliation: '1'
  - name: Steven Cannon
    affiliation: '1'
  - name: Darwin Campbell
    affiliation: '1'
  - name: Parnal Joshi
    affiliation: '1'
  - name: Iddo Friedberg
    affiliation: '1,4'
  - name: Kokulapalan Wimalanathan
    affiliation: '1,2'
  - name: Carolyn J. Lawrence-Dill
    affiliation: '1,2,3'
    email: triffid@iastate.edu
    institution: Roy J. Carver Co-Lab, Iowa State University
    street: 1111 WOI Rd
    city: Ames
    state: State IA # only USA, Australia, Canada
    zip: 50011
    country: USA
affiliation:
  - id: '1'
    department: Bioinformatics and Computational Biology
    institution: Iowa State University
    city: Ames
    state: IA # only USA, Australia, Canada
    country: USA
  - id: '2'
    department: Department of Genetics, Development, and Cell Biology
    institution: Iowa State University
    city: Ames
    state: IA # only USA, Australia, Canada
    country: USA
  - id: '3'
    department: Department of Agronomy
    institution: Iowa State University
    city: Ames
    state: IA # only USA, Australia, Canada
    country: USA
  - id: '4'
    department: Department of Veterinary Microbiology
    institution: Iowa State University
    city: Ames
    state: IA # only USA, Australia, Canada
    country: USA
date: "Febcember 32nd 3023"
language: en-US
site: "bookdown::bookdown_site"
csl: frontiers.csl
link-citations: true
output:
  bookdown::pdf_book:
    keep_tex: true
    keep_md: true
    base_format: rticles::frontiers_article
    citation_package: natbib
    includes:
      in_header:
        preamble.tex
---

```{r load_libs, include=F}
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo=FALSE)
```

```{r helper-functions}
format_genome <- function(genomes) {
  return(unlist(lapply(strsplit(genomes, "\\."), function(x) { 
    if(length(x) > 1) {
      paste0("\\textit{",gsub("_", " ",x[1]),"} ", paste(x[2:length(x)], collapse="."))
    } else {
      paste0("\\textit{",gsub("_", " ",x[1]),"} ")
    }
  }))
  )
}
```

\begin{abstract}

Abstract length and content varies depending on article type. Refer to 
\url{http://www.frontiersin.org/about/AuthorGuidelines} for abstract requirement
and length according to article type.

%All article types: you may provide up to 8 keywords; at least 5 are mandatory.
\tiny
 \keyFont{ \section{Keywords:} Text Text Text Text Text Text Text Text }

\end{abstract}

# Introduction
Hello, how are we doing?